\documentclass{article}

% "preprint" option is used for arXiv or other preprint submissions
\usepackage[preprint]{neurips_2025}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{dsfont}         % extra math symbols (indicator function)
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{graphicx}       % pictures
\usepackage{siunitx}

\usepackage{float}
\sisetup{
    round-mode          = places,
    round-precision     = 6,
    table-number-alignment = center
}

% Additional packages (personal preference)
\usepackage{amsmath, amsfonts}
\usepackage{subcaption}
\usepackage{multirow}

% Note. For the workshop paper template, both \title{} and \workshoptitle{} are required, with the former indicating the paper title shown in the title and the latter indicating the workshop title displayed in the footnote.
\title{ \Large{ICCAD 2026 CAD Contest Problem D} \\ \LARGE{Timing Fixing by Useful Skew}}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{%
  Aaron~Wu \\
  Department of Electrical Engineering\\
  National Taiwan University\\
  \texttt{b13901011@ntu.edu.tw} \\
  \And
  Yucheng~Liu \\
  Department of Electrical Engineering\\
  National Taiwan University\\
  \texttt{b13901104@ntu.edu.tw} \\
  \And
  Sheng-Yu~Cheng \\
  Department of Electrical Engineering\\
  National Taiwan University\\
  \texttt{b13901088@ntu.edu.tw} \\
  \And
  Hong-Bin~Lai \\
  Department of Electrical Engineering\\
  National Taiwan University\\
  \texttt{b13901078@ntu.edu.tw}
}

\begin{document}
%%%%%%%%%
% for consistency, We have to unify the synonyms as follows:
% agent = model = policy
%%%%%
\maketitle

\begin{abstract}
  % Briefly summarize the problem, the useful-skew optimization approach, the
  % implemented candidate/acceptance policy framework, and the main experimental
  % finding. Keep this to one compact paragraph for a 6--8 page report.
\end{abstract}

\section{Introduction}
Timing closure is a critical task in modern VLSI design. As clock frequency increases, setup and hold violations become harder to fix without disturbing the functional logic of the circuit. ICCAD Contest 2026 Problem D focuses on this setting: given an initial clock tree, a buffer library, and data-path delay reports, the goal is to improve timing quality by modifying the clock tree while preserving the original data-path logic.

The key idea behind this problem is useful skew. Instead of treating all clock skew as harmful, useful-skew optimization intentionally changes clock arrival times at flip-flops to improve timing margins. For a setup-critical path, delaying the capture clock can provide more time for data propagation. For a hold-critical path, delaying the launch clock can postpone data launch and reduce the risk of early data arrival. Therefore, by inserting or resizing buffers in the clock tree, the optimizer can improve setup and hold slack without changing combinational logic.

However, this optimization is highly coupled. A clock-tree edit does not affect only one data path; it changes the clock arrival time of an entire downstream subtree. As a result, a move that improves one violating path may degrade other paths that share the same clock-tree nodes. In addition, buffer insertion and upsizing may increase area, while aggressive area recovery may undo previous timing improvements. The objective is therefore not simply to fix the worst violation, but to balance SS setup improvement, FF hold improvement, and clock-tree area.

In this work, we formulate timing fixing by useful skew as a constrained local-search problem over legal clock-tree states. Our solver maintains a mutable clock-tree representation, evaluates timing and area metrics after local edits, and uses reversible trial operations to explore the search space efficiently. Candidate-generation policies decide which clock-tree edits should be evaluated, while acceptance strategies decide how the solver moves from one tree state to another. The experimental discussion focuses on how these policies trade off setup slack, hold slack, and area.

The rest of this report is organized as follows. Section 2 formulates the timing model, objective, and legal action space. Section 3 describes the local-search framework, including candidate generation, action evaluation, acceptance strategies, and legality checks. Section 4 reports experimental results and discusses the trade-offs among the tested optimizers. Section 5 concludes the report and outlines future improvements.


\section{Problem Formulation}

\subsection{Input Design and Legal Solution Space}

Each testcase consists of an initial clock tree, a buffer library, and two data-path delay reports under the SS and FF corners. The clock tree contains a clock source as the root, clock buffers as internal nodes, and flip-flops as leaf nodes. The buffer library specifies the available buffer types, including their area, maximum fanout constraint, and fanout-dependent delay under each timing corner. The delay reports give the clock period and the fixed data-path delay between launch and capture flip-flops.

The optimizer is allowed to improve timing only by modifying the clock tree. The data-path delay is fixed, so the combinational logic is not part of the optimization space. A legal output must preserve all original contest nodes and their names, and the relative order among original nodes in the clock tree must remain unchanged. Newly inserted buffers must have unique names, all pins must be connected, each flip-flop must have a single driver, and every buffer must satisfy the fanout limit specified by the buffer library.

We model a solution as a clock-tree state. A state records the cell type of each buffer, the parent-child connectivity of the clock tree, and the set of buffers inserted by the optimizer. The legal solution space consists of all clock trees reachable from the input tree by legal clock-tree edits while satisfying the structural constraints above.

\subsection{Clock Arrival Time and Useful Skew}
For each flip-flop, the clock arrival time is the total delay from the clock source to that flip-flop through the clock tree. Since buffer delay depends on the selected cell type, the fanout count, and the timing corner, the clock arrival time is also corner-dependent.

A data path connects a launch flip-flop to a capture flip-flop through combinational logic. Let $D_{clk}^{\mathrm{launch}}$ and $D_{clk}^{\mathrm{capture}}$ denote the clock arrival times from the clock source to the launch and capture flip-flops. We define clock skew as
\begin{equation}
\mathrm{Skew} = D_{clk}^{\mathrm{capture}} - D_{clk}^{\mathrm{launch}}.
\end{equation}
With this convention, positive skew means that the capture clock arrives later than the launch clock. A clock-tree edit may change the arrival time of an entire downstream subtree, so the skew of many data paths can change after a single local modification.
\subsection{Setup/Hold Slack and Timing Metrics}

Let $D_{\mathrm{data}}$ be the data-path delay, $T_{\mathrm{clk}}$ be the clock period, $T_{\mathrm{setup}}$ be the setup requirement, and $T_{\mathrm{hold}}$ be the hold requirement.  With the skew convention above, the setup and hold slacks are
\begin{align}
  \mathrm{Slack}_{\mathrm{setup}}
  &= T_{\mathrm{clk}} - T_{\mathrm{setup}} - D_{\mathrm{data}} + \mathrm{Skew}, \\
  \mathrm{Slack}_{\mathrm{hold}}
  &= D_{\mathrm{data}} - T_{\mathrm{hold}} - \mathrm{Skew}.
\end{align}

A negative setup slack means that data arrives too late at the capture flip-flop.  A negative hold slack means that data arrives too early and may overwrite the previous value.  Therefore, a setup violation can often be improved by increasing the capture clock delay, while a hold violation can often be improved by increasing the launch clock delay.

Timing quality is measured using total negative slack (TNS) and worst negative slack (WNS).  WNS captures the most severe timing violation among all paths, while TNS measures the accumulated amount of negative slack.  In this problem, setup timing is evaluated under the setup-oriented SS corner, and hold timing is evaluated under the hold-oriented FF corner.

\subsection{Timing-Area Objective and Internal Proxy Score}
The objective combines timing and area.  Timing quality is measured by TNS and WNS under the SS and FF corners, while area is the total area of the clock-tree buffers.  Given a baseline metric value $x^{\mathrm{ori}}$ and a final value $x$, we use the relative improvement form
\begin{equation}
I(x, x^{\mathrm{ori}}) = 1 - \frac{x}{x^{\mathrm{ori}}},
\end{equation}
when $x^{\mathrm{ori}}$ is nonzero.  This form rewards reducing negative timing metrics toward zero and rewards reducing area below the baseline.

The official contest score is a weighted sum of setup improvement, hold improvement, and area reduction:

\begin{equation}
  \begin{aligned} \mathrm{Score} &= \alpha \left[ \left( 1 - \frac{\mathrm{TNS}_{SS}}{\mathrm{TNS}_{SS}^{ori}} \right) + \left( 1 - \frac{\mathrm{WNS}_{SS}}{\mathrm{WNS}_{SS}^{ori}} \right) \right] + \beta \left[ \left( 1 - \frac{\mathrm{TNS}_{FF}}{\mathrm{TNS}_{FF}^{ori}} \right) + \left( 1 - \frac{\mathrm{WNS}_{FF}}{\mathrm{WNS}_{FF}^{ori}} \right) \right] \\ &+ \gamma \left( 1 - \frac{\mathrm{Area}}{\mathrm{Area}^{ori}} \right).
  \end{aligned}
\end{equation}

The official weights are not disclosed, so the optimizer cannot directly use the official score as its search objective.  According to the Q\&A session, the organizers indicated that the weights roughly satisfy $\alpha > \beta \approx \gamma$.  Based on this hint, we use a proxy score with fixed weights $\alpha = 0.5$, $\beta = 0.25$, and $\gamma = 0.25$ to guide move selection during optimization.  These parameters are not claimed to be the hidden official weights; they are used only as an internal ranking function that reflects the relative importance of setup repair, hold repair, and area control.

This objective captures the main trade-off in the problem.  Aggressive buffer insertion or upsizing may improve setup or hold slack, but the same move can reduce the area component and may also hurt timing on other paths.  On the other hand, area recovery may remove unnecessary delay but can also undo a previous timing repair.  Therefore, the optimizer must balance timing improvement and area cost rather than optimizing only one metric independently.

\subsection{Local Edit Semantics}

Our optimizer explores the solution space using three local edit primitives during search.  First, an \emph{insert} move places a new buffer on an existing clock-tree edge.  This increases the clock delay of the downstream subtree and can be used to adjust useful skew, but it also increases area.

Second, a \emph{resize} move changes the cell type of an existing buffer. Resizing gives a finer way to increase or decrease clock delay and area without changing the clock-tree topology.  Depending on the selected cell type, a resize move may either improve timing by adding delay or reduce area by using a smaller buffer.

Third, a \emph{remove} move deletes a buffer that was inserted by the optimizer in an earlier step and reconnects the surrounding tree.  This move is used only as an internal search operation for reversible exploration and area recovery. Original contest nodes remain part of the design and are never removed.  Thus, the final output still satisfies the legal modification constraints of the problem.

Because the number of possible edit sequences is large, the solver cannot enumerate the full state space.  Instead, each iteration generates a smaller set of candidate moves and evaluates their impact on the current timing and proxy score.  The rest of the report describes how we construct these candidate sets and how different search policies decide which move to accept.

\section{Methodology}

\subsection{Solver Overview}

After defining the legal edit operations and the timing-area objective, the remaining question is how to search the large space of legal clock-tree states efficiently. Our solver uses a local-search framework that repeatedly moves from the current clock tree to a nearby legal state. Each iteration is organized around two decisions: which candidate moves should be evaluated, and which evaluated move should be accepted as the next state.

We separate these two decisions into a \textbf{candidate-generation policy} and an \textbf{acceptance strategy}. The candidate-generation policy defines the local neighborhood visible to the solver in the current iteration. It may focus on violating timing paths, expand around upstream clock-tree regions, or include remove and resize moves for area recovery. The acceptance strategy then decides how the solver moves inside this neighborhood. It may greedily accept the best immediate improvement, probabilistically accept a temporary degradation, or use short-term memory to avoid cycling.

Algorithm~\ref{alg:local-search} summarizes the overall optimization flow. Each iteration first constructs a bounded candidate set, evaluates the candidates through temporary application, and then applies one move according to the acceptance strategy. The solver also maintains the best-seen state separately from the current state, because exploratory strategies may temporarily move to lower-scoring states.

\begin{algorithm}[htbp]
  \caption{Overall local-search optimization flow}
  \label{alg:local-search}
  \begin{algorithmic}[1]
    \State Build the initial \texttt{ClockTree}, \texttt{DataPathGraph}, and \texttt{TimingState}
    \State Evaluate the baseline tree and compute its proxy score
\State $T_{\mathrm{cur}} \gets$ baseline clock tree
\State $T_{\mathrm{best}} \gets T_{\mathrm{cur}}$
\State $S_{\mathrm{best}} \gets S_{\mathrm{proxy}}(T_{\mathrm{cur}})$
\While{current time $<$ time budget}
\State Generate a candidate move set $M$ using the candidate policy
\State $E \gets \emptyset$
\ForAll{$m \in M$}
\State Temporarily apply $m$ to $T_{\mathrm{cur}}$
\State Update timing and area metrics
\State Compute $\Delta S_{\mathrm{proxy}}(m)$
\State Store the evaluated move information in $E$
\State Undo $m$ and restore $T_{\mathrm{cur}}$
\EndFor
\State Select a move $m^\star$ from $E$ using the acceptance strategy
\If{$m^\star$ exists}
\State Apply $m^\star$ to $T_{\mathrm{cur}}$
\State Update \texttt{TimingState} for the current tree
\If{$S_{\mathrm{proxy}}(T_{\mathrm{cur}}) > S_{\mathrm{best}}$}
\State $T_{\mathrm{best}} \gets T_{\mathrm{cur}}$
\State $S_{\mathrm{best}} \gets S_{\mathrm{proxy}}(T_{\mathrm{cur}})$
\EndIf
\EndIf
\EndWhile
\State Restore $T_{\mathrm{best}}$
\State Perform full timing evaluation
\State Write \texttt{modified\_clk\_tree.structure}
\end{algorithmic}
\end{algorithm}

The rest of this section follows the structure of Algorithm~\ref{alg:local-search}. Section~3.2 describes the optimization state and the incremental timing evaluation used inside the temporary apply-score-undo loop. Section~3.3 describes how candidate-generation policies construct the move set $M$. Section~3.4 describes how acceptance strategies choose the move $m^\star$ from the evaluated set. Finally, Section~3.5 explains how the best-seen internal state is converted into a legal output file.
\subsection{Optimization State and Incremental Evaluation}

The optimizer maintains three main data structures during search. The \texttt{ClockTree} is the mutable solution state and stores the active topology and buffer cell types. The \texttt{DataPathGraph} stores the fixed timing paths, including the launch flip-flop, capture flip-flop, and data delay of each path. Since useful-skew optimization modifies only the clock tree, this graph remains read-only. The \texttt{TimingState} connects the two structures and caches the quantities needed for fast evaluation, including clock arrivals, setup and hold slacks, TNS, WNS, area, and the proxy score.

Candidate evaluation follows the temporary apply-score-undo loop shown in Algorithm~\ref{alg:local-search}. For each trial edit, the solver applies the move to the \texttt{ClockTree}, updates the affected clock arrivals and related timing paths in \texttt{TimingState}, and computes the proxy-score change. The move is then either accepted or rolled back. Rollback restores both the clock-tree state and the timing cache, so rejected candidates do not affect later evaluations.

This design avoids recomputing all clock arrivals and slacks for every trial move. Instead, the solver updates only the parts of the timing state affected by the local edit, allowing many candidate moves to be evaluated within the fixed time budget.

\subsection{Candidate Generation}

Since the legal edit space is too large to evaluate exhaustively, candidate generation defines the local neighborhood seen by the optimizer at each iteration. A good candidate policy should include moves that are likely to improve useful skew, but it should also keep the candidate set small enough that many iterations can be completed within the time budget. We therefore explored several policies with different degrees of timing awareness and search breadth.

The simplest policy is \textbf{random generation}. It uniformly samples legal insert, remove, and resize moves from the current clock tree. This policy is the most naive way to explore the solution space, and it does not directly use timing information to guide move selection. As a result, it may generate many weak moves that do not improve timing. However, it can also expose non-obvious area-recovery or resizing opportunities that more focused policies may miss.

The most direct timing-driven policy is \textbf{violation path generation}. It scans all data paths, selects paths with negative slack, and prioritizes the most severe violations. For setup violations, it generates moves near the capture flip-flop; for hold violations, it generates moves near the launch flip-flop. This follows the useful-skew formulas directly, but the resulting neighborhood can be too narrow when the best repair is not close to the violating endpoint.

To address this limitation, the \textbf{upstream-window policy} expands the search region from a violating endpoint toward the root of the clock tree. Instead of considering only the endpoint edge, it considers several upstream buffers and edges. These moves may affect a larger downstream subtree and can therefore repair several related timing paths at once.

The \textbf{critical-endpoint policy} uses path violations to assign a criticality value to each flip-flop. Negative setup slack contributes to the criticality of the capture flip-flop, while negative hold slack contributes to the criticality of the launch flip-flop. The policy then generates moves around the most critical flip-flops. Compared with selecting individual violating paths, this policy tries to identify endpoints that participate in many violations and may therefore produce moves with broader timing impact.

Finally, the \textbf{union-pool policy} combines the action sources of violation path, upstream window, and critical endpoint into one candidate pool. It uses the same total sampling budget as the individual policies, so the goal is not to evaluate more candidates, but to obtain a more diverse candidate set under the same cost. For simulated-annealing-based optimizers, we use a \textbf{sampled-union-pool policy}, which samples only one move from this mixed pool in each iteration to match the single-proposal style of simulated annealing.


\subsection{Acceptance Strategies}

After candidate generation defines the visible neighborhood, the acceptance strategy determines how the search moves through that neighborhood. This decision is important because the proxy score is not a smooth objective. A move that is locally worse may still help the solver reach a better region later, while a purely greedy sequence may quickly become trapped after all immediate improvements are exhausted.

The \textbf{greedy strategy} is the most exploitative choice. It evaluates all candidates in the current set, and applies the move with the largest positive proxy-score improvement. This strategy is stable and easy to interpret: every accepted move immediately improves the current solution according to the proxy score. However, its performance depends heavily on the candidate pool. If the pool does not contain a good improving move, greedy search cannot escape the current local optimum.

The \textbf{two-step strategy} is a greedy strategy with two objective phases. The first phase greedily accepts moves that improve a slack-focused objective, with stronger emphasis on setup TNS and WNS. The second phase switches back to the full proxy score, so setup, hold, and area are optimized together. This design tries to create timing margin first and then recover overall score, but Its weakness is that area recovery may undo previous timing repairs, and timing-focused moves may create too much area overhead for the second phase to repair.

\textbf{Simulated annealing}\cite{kirkpatrick1983optimization} introduces stochastic exploration. At each step, it samples a proposal, evaluates its proxy-score change, and always accepts improving moves. Different from greedy search, a worsening move may also be accepted with a temperature-controlled probability. At high temperature, the search can move through worse intermediate states and escape local optima; at low temperature, it behaves more like greedy search. The solver still records the best-seen state separately, so temporary degradation does not directly determine the final output.

\textbf{Iterated simulated annealing} repeats this exploration process in several rounds. Between rounds, the solver restores the best-seen state and performs small greedy cleanup batches. The SA phases provide exploration, while the cleanup phases exploit local improvements around promising regions. This design attempts to keep the ability to escape local optima without letting the current state drift too far from a high-quality solution.

\textbf{Tabu search} \cite{glover1989tabu} uses short-term memory instead of random temperature to guide exploration. At each step, it evaluates the candidate set and chooses the best non-tabu move. Recently used moves are marked tabu so that the solver is discouraged from cycling between the same few states. A tabu move can still be accepted by aspiration if it improves the global best score. This makes tabu search more exploratory than greedy search while still keeping each step strongly score-driven.

Overall, candidate generation and acceptance play complementary roles. Candidate policies decide which legal edits are worth looking at, while acceptance strategies decide how aggressively the solver should exploit or explore the evaluated moves. The optimizer variants evaluated in Section 4 are different combinations of these two choices.

\subsection{Output Format Validation}

Output validity is enforced during move construction rather than repaired after optimization. Inserted buffers receive unique generated names, resize and insertion moves are rejected if the selected cell type violates the fanout limit, and remove moves are limited to optimizer-inserted buffers. Original contest nodes remain active and keep their names, while removed search-only buffers are marked inactive so that they cannot appear in the final output.

After the optimizer finishes, the solver restores the best-seen \texttt{ClockTree} and writes only active nodes through a preorder traversal with depth information, matching the input clock-tree format. Since the writer emits only active parent-child relationships, each flip-flop keeps a single clock-tree parent, the tree remains well formed, and every emitted buffer satisfies the buffer-library fanout table. The final file is written through a temporary file followed by a rename, so the output path is replaced only after a complete \texttt{modified\_clk\_tree.structure} file has been produced.


\section{Experimental Results and Discussion}
\label{sec:experiments}

\subsection{Experimental Setup and Metrics}

We evaluate all optimizer variants on five public testcases.  Each optimizer is
run with five random seeds, giving $13 \times 5 \times 5 = 325$ completed runs.
The reported score is the proxy score defined in Section~2, using
$\alpha=0.5$, $\beta=0.25$, and $\gamma=0.25$.

In the result tables, Setup and Hold denote $S_{\mathrm{setup}}$ and
$S_{\mathrm{hold}}$, i.e., the sum of TNS and WNS improvement for the SS and FF
corners respectively.  Area denotes $S_{\mathrm{area}}$, where a larger value
means lower final buffer area relative to the baseline.

\subsection{Effect of Candidate Generation}

We first isolate the effect of candidate generation by fixing the acceptance
strategy to greedy best-score selection.  Under this setting, the optimizer can
only accept an immediately improving move, so the result mainly reflects whether
the candidate pool contains useful local edits.

Union pool performs best among the greedy variants, while random generation and
upstream window are close behind.  This suggests that greedy search benefits
from a more diverse candidate set under the same sampling budget.  Violation
path and critical endpoint are weaker because their candidate regions are too
narrow: they focus on directly violated endpoints, but useful clock-tree edits
often appear farther upstream or in resize/remove moves that help area recovery.

\subsection{Effect of Acceptance Strategy}

We next fix the candidate family and compare how different acceptance strategies
move through the generated neighborhood.  This experiment separates the quality
of candidate generation from the ability of the search strategy to escape local
optima.

\subsubsection{Random Candidate Policy}

With random candidates, tabu search gives the best result, while two-step
optimization is the weakest.  Random candidates provide a broad but noisy search
space.  Tabu search can exploit this space because it still evaluates candidate
quality, avoids recent cycles through tabu memory, and restores the best
solution found.  In contrast, the two-step method separates timing repair and
area recovery too rigidly.  Timing-focused moves may leave too much area cost
for the second stage to clean up, while area-recovery moves may undo previous
timing improvements.  Its timing-first stage also does not outperform SA/ISA
enough to justify this cleanup cost.

\subsubsection{Union-Based Candidate Policies}

With union pool and sampled union pool, two-step optimization remains weak.
Although union pool provides a more diverse candidate set under the same budget,
the fixed two-stage schedule does not convert this diversity into a better final
score.  SA and ISA aggressively improve setup slack from sampled union
proposals, but this improvement often comes with a large area penalty because
many accepted moves insert or preserve extra buffers.  Tabu search is more
balanced: it improves timing while keeping final area close to the baseline.

\subsection{Overall Optimizer Comparison}

Table~\ref{tab:optimizer_results} summarizes the final score decomposition for
all optimizer variants.  This table combines the two ablations above and shows
which policy combination gives the best overall timing-area trade-off.

\begin{table}[H]
\centering
\caption{Performance comparison of different optimizers. Std. is computed from
the final scores over five testcases and five seeds.}
\label{tab:optimizer_results}

\begin{tabular}{
l
S[table-format=1.6]
S[table-format=1.6]
S[table-format=1.6]
S[table-format=-1.6]
S[table-format=1.6]
}
\toprule
{Optimizer} &
{Avg. Score} &
{Setup} &
{Hold} &
{Area} &
{Std.} \\
\midrule

tabu-random              & \bfseries 1.140722 & 1.276422 & \bfseries 1.916255 &  0.093788 & 0.235526 \\
isa-random               & 1.084773 & 1.322733 & 1.737303 & -0.043676 & 0.273375 \\
tabu-union-pool          & 1.062058 & 1.259221 & 1.731261 & -0.001474 & 0.274410 \\
isa-sampled-union-pool   & 1.044759 & 1.521136 & 1.721173 & -0.584406 & 0.183581 \\
sa-sampled-union-pool    & 1.044015 & 1.653149 & 1.833949 & -0.964191 & \bfseries 0.165838 \\
greedy-union-pool        & 1.042167 & 1.193238 & 1.757868 &  0.024327 & 0.266567 \\
greedy-upstream-window   & 1.031489 & 1.227970 & 1.710767 & -0.040751 & 0.275511 \\
greedy-random            & 1.009059 & 1.104560 & 1.806695 &  0.020420 & 0.214807 \\
sa-random                & 0.998808 & 1.149864 & 1.750100 & -0.054596 & 0.276991 \\
two-step-random          & 0.876061 & 1.148869 & 1.334986 & -0.128479 & 0.318905 \\
two-step-union-pool      & 0.859649 & 1.045729 & 1.405060 & -0.057920 & 0.401216 \\
greedy-violation-path    & 0.829324 & 1.032678 & 1.300464 & -0.048521 & 0.277805 \\
greedy-critical-endpoint & 0.746583 & 0.923081 & 1.188662 & -0.048491 & 0.375666 \\

\bottomrule
\end{tabular}
\end{table}

\textit{Tabu-random} achieves the highest average score.  It does not have the
largest setup component, but it has the best hold component and also reduces
area.  In contrast, \textit{sa-sampled-union-pool} has the highest setup
component, but its strongly negative area score pulls down its final score.
This shows that maximizing one timing component is not sufficient: the final
ranking is determined by the balance among setup repair, hold repair, and area
control.

\subsection{Discussion: Why Tabu-Random Performs Best}

The final result can be explained by the interaction between candidate diversity
and search memory.  SA and ISA are able to escape local optima, and sampled
union proposals make them strong at setup repair.  However, their stochastic
acceptance can keep many area-increasing moves, so their timing gain is partly
lost through area penalty.  Greedy search has the opposite behavior: it is stable
and area-aware, but it relies heavily on the candidate pool and cannot move
through temporary score degradation.

Tabu-random gives a better compromise.  Random candidates expose non-local
insert, remove, and resize opportunities that a violation-focused union pool may
miss.  Tabu memory then prevents the search from repeatedly returning to recent
moves while still choosing the best non-tabu candidate at each step.  This
explains why \textit{tabu-random} outperforms \textit{tabu-union-pool}: the
random neighborhood is less biased toward immediate violation repair, and tabu
search is strong enough to extract useful moves from that broader neighborhood.
As a result, \textit{tabu-random} does not maximize setup alone, but it finds
the best overall setup-hold-area balance.


\section{Conclusion}
This work formulates useful-skew timing fixing as a constrained local-search
problem over legal clock-tree states.  The solver modifies only the clock tree
through buffer insertion, inserted-buffer removal, and buffer resizing, and uses
a proxy score to balance SS setup repair, FF hold repair, and area control.
Internally, a mutable \texttt{ClockTree}, a fixed \texttt{DataPathGraph}, and an
incremental \texttt{TimingState} support reversible candidate evaluation, so
different search policies can be compared while keeping the final output legal.

The experimental results show that both candidate generation and acceptance
strategy are decisive.  Greedy search is stable and benefits from the diverse
union-pool neighborhood, but it cannot pass through temporary score degradation.
SA and ISA explore more aggressively and achieve strong setup improvement with
sampled-union proposals, but their area penalty reduces the final score.
Two-step optimization captures the intended timing-first and cleanup structure,
yet the fixed phase boundary is too rigid for the coupled setup-hold-area
objective.

Among all tested optimizers, \textit{tabu-random} achieves the best average
score.  Its advantage comes from combining a broad random neighborhood with
score-driven tabu memory: random candidates expose non-local insert, remove, and
resize opportunities, while tabu search avoids short cycles and keeps the
current trajectory close to high-quality states.  As a result, it does not
maximize setup improvement alone, but obtains the best balance among setup
repair, hold repair, and area reduction.  Future work could improve this
framework by adapting the candidate policy to the current violation pattern,
calibrating the proxy weights with more feedback, and adding a stronger
late-stage area-aware resize/remove polish.

\section*{Author Contributions}
Hong Bin Lai did nothing :\(\)) That's True, I'm very sorry.I find the branch yesterday but I see the first version submission when I'm writing.I know it's a bad excuse, and I really doesn't do anything.

\section*{Generative AI Usage Disclosure}
We used generative AI tools during the development of this project. In particular, OpenAI Codex and Cursor was used extensively to assist with source-code implementation, including parser construction, clock-tree data structures, timing evaluation routines, optimization-loop implementation, test scripts, and refactoring. Approximately 80\% of the submitted source code was initially generated or modified with the assistance of Codex or Cursor.

The algorithmic ideas, including the useful-skew optimization strategy, candidate generation policies, acceptance strategies, timing-area trade-off considerations, experimental design, and final optimizer selection, were designed and evaluated by the team members. All AI-generated code was manually reviewed, integrated, debugged, and tested by the team. The reported experimental results were obtained by executing our implemented solver on the test cases; no experimental data or performance numbers were generated by AI tools.

We also used ChatGPT/Codex to assist with report organization, wording refinement, and documentation polishing. The final technical content, claims, results, and conclusions were checked and approved by the team.


\medskip
{
  \small
  \bibliographystyle{unsrt}
  \bibliography{references}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}
